---
title: "k-NN Algorithm"
output: html_document
date: "2024-06-22"
---

```{r setup, include=FALSE}
# Install prerequisite packages if not already installed
if (!requireNamespace("dplyr", quietly = TRUE)) {
  install.packages("dplyr", dependencies = TRUE)
}
if (!requireNamespace("ggplot2", quietly = TRUE)) {
  install.packages("ggplot2")
}
if (!requireNamespace("ggfortify", quietly = TRUE)) {
  install.packages("ggfortify")
}

# Load required libraries
library(dplyr)
library(ggplot2)
library(ggfortify)
```

##Reading the Data Set
```{r}

diabetes_data8173= read.csv("~/Documents/R Program/diabetes.csv")
diabetes_data8173
```

## Summarizing the dataset
```{r}
summaryDiabetesData_8173<-summary(diabetes_data8173)
summaryDiabetesData_8173
```

```{r pressure, echo=FALSE}
plot(pressure)
```

## Showcasing the structure and dimension of the dataset
```{r}
str(diabetes_data8173)
dim(diabetes_data8173)
```

## Showing the class variable in diabetes dataset?

```{r}
class(diabetes_data8173)
```

# Changing the class type of the class variable of diabetes dataset to factor. 
#Showing the output after the conversion.


```{r}
diabetes_data8173$outcome <- as.factor(diabetes_data8173$outcome) 
print("Outcome column after conversion:")
str(diabetes_data8173$outcome)
```
```{r}
colSums(is.na(diabetes_data8173))
```

# Replace the missing values in diabetes by 0. 
```{r}
diabetes_data8173[(is.na(diabetes_data8173))]<-0
print("After replacing missing values with 0:")
colSums(is.na(diabetes_data8173))
```
# preparing data for plotting by removing outcome column
```{r}
plotDiabetes_data8173=select(diabetes_data8173,c(1,2,3,4,5,6,7,8))
names(diabetes_data8173)
```

# Perform k-means clustering
```{r}
```


```{r}
kmeans.diabetesResult2_8173 <- kmeans(plotDiabetes_data8173, 2)
```

# Plotting the data
```{r}
plot(plotDiabetes_data8173[c("plas", "age")], col = kmeans.diabetesResult2_8173$cluster)
```
```{r}
kmeans.diabetesResult3_8173 <- kmeans(plotDiabetes_data8173, 3)
# Define custom colors for clusters
cluster_colors <- c("red", "green", "orange")
#plotting the data
plot(plotDiabetes_data8173[c("preg", "insu")], col = cluster_colors[kmeans.diabetesResult3_8173$cluster],
     main = "K-Means Clustering of Diabetes Data",
     xlab = "Blood Pressure (pres)",
     ylab = "Insulin (insu)")
legend("topright", legend = unique(kmeans.diabetesResult3_8173$cluster), col = cluster_colors, pch = 1)
```


#Plotting
```{r}
kmeans.diabetesResult4_8173 <- kmeans(plotDiabetes_data8173, 4)
# Define custom colors for clusters
cluster_colors <- c("blue","violet", "cyan", "green")
#plotting the data
plot(plotDiabetes_data8173[c("pres", "insu")], col = cluster_colors[kmeans.diabetesResult4_8173$cluster],
     main = "Scatter plot of Diabetes Data",
     xlab = "Blood Pressure (pres)",
     ylab = "Insulin (insu)")

plot(plotDiabetes_data8173, col = kmeans.diabetesResult3_8173$cluster)
```

